A Survey of Sterile Neutrinos

### 2015-04-14 p3_505/paper/paper3.tex
doug updates to paper

inside of a large Cherenkov detector filled with liquid scintillator, utilizing 1280 phototubes. Figure~\ref{fig:lsnd_miniboone}(L) shows the excess of $\overline{\nu}_e$ events observed by LSND. The observation was $87.9\pm 22.4\pm6$ events over expected background. Their final result claimed the existence of $\overline{\nu}_\mu\rightarrow\overline{\nu}_e$ oscillations at $\Delta m^2 \sim 1\text{ eV}$ with 3.8$\sigma$. This is in heavy conflict with the existing three light flavor neutrino model supported by atmospheric and solar neutrino oscillation experiments, as well as measurements at LEP predicting $N=2.984\pm0.0082$ for weakly interacting neutrino flavors. Therefore, at least one ``sterile'' neutrino is required.

@@

  \caption{(L) From the LSND Experiment~\cite{LSND}. The expected number of events is represented by the red and green histograms, the blue histogram agrees with the data and models $\overline{\nu}_{\mu}\rightarrow \overline{\nu}_e$ oscillation. (C) The MiniBooNE electron event excess in neutrino and anti-neutrino sector~\cite{mini2}. (R) The MiniBooNE allowed regions in neutrino and anti neutrino mode, for a 2 neutrino model. Black stars represent the MiniBooNE best fit points for $\Delta m^2$ and $\sin^22\theta$.~\cite{mini2}}

@@

The full dataset from MiniBooNE was consistent with the LSND $\overline{\nu}_e$ excess, and therefore the sterile neutrino hypothesis survives with a best fit $\Delta m^2 \sim 1$ eV.
\subsection{Reactor Anomaly}
reactors are cool.

\section{Current and Future Experimental Efforts}
Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
\subsection{Current Experiments}
%Matt
\subsubsection{MINOS/MINOS+}
The MINOS (Main Injector Neutrino Oscillation Search) experiment ran from 2005 to 2012 and its wide-band fresh, MINOS+, has taking data since September 2013. The MINOS experiment is a long baseline style detector with a neutrino beam source and near detector at Fermilab and a far detector located in northern Minnesota. Though its primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data. MINOS detects sterile neutrinos via their disappearance in neutral-current (NC) interactions, which are independent of the three known neutrino flavors and their oscillations, and in charged-current (CC) interactions differing from those predicted from the mixing of the three neutrinos of the standard model. The latest sterile neutrino results from the collaboration \cite{MINOS} now exclude a large majority of the LSND allowed phase space, Figure~\ref{fig:MINOS}. These exclusion limits are predicted to increase by up to a factor of 2 as MINOS+ continues to run.

@@

\subsection{Future Experiments}
%Matt
NOVA
MicroBooNE
DUNE (Formerly LBNE)

>>>>

inside of a large Cherenkov detector filled with liquid scintillator, utilizing 1280 phototubes. Figure~\ref{fig:lsnd_miniboone}L shows the excess of $\overline{\nu}_e$ events observed by LSND. The observation was $87.9\pm 22.4\pm6$ events over expected background. Their final result claimed the existence of $\overline{\nu}_\mu\rightarrow\overline{\nu}_e$ oscillations at $\Delta m^2 \sim 1\text{ eV}$ with 3.8$\sigma$. This is in heavy conflict with the existing three light flavor neutrino model supported by atmospheric and solar neutrino oscillation experiments, as well as measurements at LEP predicting $N=2.984\pm0.0082$ for weakly interacting neutrino flavors. Therefore, at least one ``sterile'' neutrino is required.

@@

  \caption{(L) From the LSND Experiment~\cite{LSND}. The expected number of events is represented by the red and green histograms, the blue histogram agrees with the data and models $\overline{\nu}_{\mu}\rightarrow \overline{\nu}_e$ oscillation. (C) The MiniBooNE electron event excess in neutrino and anti-neutrino sector~\cite{mini2}. (R) The MiniBooNE allowed regions in neutrino and anti neutrino mode, for a 2 neutrino model. Black stars represent the MiniBooNE best fit points for $\Delta m^2$ and $\sin^22\theta$~\cite{mini2}.}

@@

The full dataset from MiniBooNE was consistent with the LSND $\overline{\nu}_e$ excess, and therefore the sterile neutrino hypothesis survived. Figures~\ref{fig:lsnd_miniboone}C and \ref{fig:lsnd_miniboone}R show the electron excess and allowed oscillation regions from the MiniBooNE data, respectively.

The systematics of the MiniBooNE detector made it difficult to differentiate between $\pi^0 \rightarrow \gamma\gamma$ and an electron in the detector; therefore, there is a large uncertainty in the total number of $\overline{\nu}_e$ events. It is also hypothesized that the neutrino flux from the Booster neutrino beam must be studied more rigorously.
\subsection{Reactor Anomaly}
There has been an effort to re-analyse short baseline reactor neutrino oscillation data using newer flux calculations. These calculations show the posibility of the disappearane of reactor $\overline{\nu}_e$.
\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{../figures/reactor_anom.png}
  \caption{Plot for the short baseline reactor antineutrino anomaly. The red curve shows a possible three active neutrino mixing solution, using $\sin^2\theta_{13} = 0.06$. The blue curve shows a possible solution including a new neutrino mass state with $\Delta m^2 \gg 1\text{ eV}^2$.}
\end{figure}

\section{Current and Future Experimental Efforts}
Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
\subsection{Current Experiments}
% Matt
\subsubsection{MINOS/MINOS+}
The MINOS (Main Injector Neutrino Oscillation Search) experiment ran from 2005 to 2012 and its wide-band fresh, MINOS+, has taking data since September 2013. The MINOS experiment is a long baseline style detector with a neutrino beam source and near detector at Fermilab and a far detector located in northern Minnesota. Though its primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data. MINOS detects sterile neutrinos via their disappearance in neutral-current (NC) interactions, which are independent of the three known neutrino flavors and their oscillations, and in charged-current (CC) interactions differing from those predicted from the mixing of the three neutrinos of the standard model. The latest sterile neutrino results from the collaboration \cite{MINOS} now exclude a large majority of the LSND allowed phase space, Figure~\ref{fig:MINOS}. These exclusion limits are predicted to increase by up to a factor of 2 as MINOS+ continues to run.

@@

\subsection{Future Experiments}
% Matt
NOVA
MicroBooNE
DUNE (Formerly LBNE)
